
%\documentclass[10pt,onecolumn,draftcls,conference]{IEEEtran}
%\documentclass[conference]{IEEEtran}
%\documentclass[preprint]{style/sigplanconf_fixed}
%\documentclass{style/sig-alternate}
%\documentclass[pageno]{style/jpaper}
%\documentclass[10pt,doublecolumn,conference]{IEEEtran}

% Format:  http://icpp2016.cs.wcupa.edu/cfp.html
% IEEEtrans without parameters is also used by our previous iCPP paper.
\documentclass[10pt, conference, compsocconf]{IEEEtran}

%replace XXX with the submission number you are given from the ASPLOS submission site.
%\newcommand{\asplossubmissionnumber}{267}

% Include my styles
\usepackage{paper}
\usepackage{color}

%\renewcommand{\baselinestretch}{0.99}
%\newcommand{\todo}[1]{\textcolor{red}{#1}}

\begin{document}

%
% paper title
%
\title{Compiler-Assisted Overlapping of Communication and Computation in MPI Applications}

\author{\IEEEauthorblockN{Jichi Guo, Qing Yi}
  \IEEEauthorblockA{University of Colorado, Colorado Springs \\
  Colorado Springs, CO, USA\\
  \{jguo2, qyi\}@uccs.edu}
  \and 
  \IEEEauthorblockN {Jiayuan Meng}
  \IEEEauthorblockA {Google Inc.  \\
  Mountain View, CA, USA \\
         meng.jiayuan@gmail.com}
  \and 
  \IEEEauthorblockN{Junchao Zhang, Pavan Balaji}
  \IEEEauthorblockA{Argonne National Laboratory \\
  Lemont, IL, USA\\
  \{jczhang, balaji\}@anl.gov}}
%\author{Jichi Guo,$^\dag$ Qing Yi,$^\dag$ Jiayuan Meng,$^\ddag$ Junchao Zhang,$^\ddag$ Pavan Balaji$^\ddag$\\
%University of Colorado Colorado Springs$^\dag$\\
%Colorado Springs, CO, USA\\
%\{jguo2, qyi\}@uccs.edu \\
%Argonne National Laboratory$^\ddag$\\
%Lemont, IL, USA\\
%meng.jiayuan@gmail.com, \{jczhang, balaji\}@anl.gov
%}
%\authorinfo{}{}{}

% PLEASE ADD INSTITUTIONS
% make the title area
\maketitle

\begin{abstract}
The performance of distributed-memory applications, many of which are
written in MPI, critically depends on how well the applications can
ameliorate the long latency of data movement by overlapping
them with ongoing computations, thereby minimizing wait time.  This
paper aims to
enable such overlapping in large MPI applications and presents a
framework that uses an analytical performance model and an optimizing compiler
to systematically enable the optimizations.  In particular, we first generate an
analytical performance model of the application execution flow to automatically identify potential communication hot spots
that may induce long wait time.  Next, for each communication hot
spot, we search the execution flow graph to find surrounding loops that
include sufficient local computation to overlap with the
communication.  Then, blocking MPI communications are decoupled into
nonblocking operations when necessary, and their surrounding loops are
transformed to hide the communication latencies behind local
computations.  We evaluated our framework using 7 MPI applications
from the NPB \texttt{NAS} benchmark suite.  Our optimizations can attain
3-88\% speedup over the original implementations.
\end{abstract}

\begin{IEEEkeywords}
Computer Applications; Computer performance; Parallel machines; Automatic programming
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

%\lstdefinelanguage{sklt}{
%  morecomment=[s]{/*}{*/},
%  morecomment=[l]{//},
%  morekeywords={
%    dummy,
%    def, call, from, import,
%    stream, for, forall, switch, case, if, else, return, break, continue,
%    st, ld, comp, ilp, blk, block,
%    int, long, float, double, char, string,
%    for_threads, grid, thread_block, commit,
%    comm, send, recv, bcast, barrier, allreduce, sendrecv, all2all, reduce,
%  },
%}
%
%\lstset{
%  language=sklt,
%  keywordstyle=\bf,
%  %keywordstyle=\em\bfseries,
%  basicstyle=\ttfamily\scriptsize,
%  numbers=left,
%  numberstyle=\tiny,
%  stepnumber=2,
%  numbersep=3pt,
%}

\input{text/intro}
\input{text/model}
\input{text/profit}
\input{text/tune}
\input{text/exp}
\input{text/related}
\input{text/conclusion}

\REMOVE{
\section*{Acknowledgment}
The research is partially funded by the National Science Foundation of
USA under Grants 1261778, 1261811, and 1261.  This research used
resources of the Argonne Leadership Computing Facility at Argonne
National Laboratory, which is supported by the Office of Science of
the U.S. Department of Energy under contract DE-AC02-06CH11357.
}


\bibliographystyle{plain}
\bibliography{bib/refs,bib/hotpath,bib/qing,bib/mpi,bib/jichi,bib/bench,bib/cco,bib/inline}
%\bibliography{venkat, jmeng, kk}

%\clearpage
\REMOVE{
\framebox{
\parbox{3.2in}{
The submitted manuscript has been created by UChicago Argonne, LLC,
Operator of Argonne National Laboratory (``Argonne"). Argonne, a
U.S. Department of Energy Office of Science laboratory, is operated
under Contract No. DE-AC02-06CH11357. The U.S. Government retains for
itself, and others acting on its behalf, a paid-up nonexclusive,
irrevocable worldwide license in said article to reproduce, prepare
derivative works, distribute copies to the public, and perform
publicly and display publicly, by or on behalf of the Government.
}
}
}

% that's all folks
\end{document}

% EOF
