\subsection{Correctness Of Optimization}
\label{sec:cco:safety}

\todo{Make an algorithm Figure. Reference the algorithm figure,
  summarize the key steps: Given the loops that containing the
  computation and communication to overlap, the safety analysis will
  check if it is safe to reorder them, and what variables are needed
  to be privatized to enable the overlap transformation.}

\todo{don't make up terms such as CCO loop, it's not even a name}
\todo{reorganize the following to following an algorithm figure}

As shown in Figure~\ref{fig:ft} , loop can be split into separation of
computation can communication.  After reordering the computation and
communication for CCO, three loop iterations $I-1$, $I$ and $I+1$ will
be \emph{interleaved}.  $After(I+1)$ will be executed before
$Communication(I)$ and $Before(I-1)$.  To guarantee the safety of the
reordering, there must be no loop-carried dependence from $After(I+1)$
to $Communication(I)$ and $Before(I-1)$.

However, in the applications such as NAS FT, there are often buffer
arrays shared by different iterations that could introduce such
loop-carried dependence.  When that happens, the buffers are needed to
be privatized to each loop iteration in order to apply CCO
optimization.  If the buffers are also used as the output of the loop,
the privatized buffers are needed to be copied to the output array.
But if the dependence from $After(I-1)$ to $Before(I+1)$ cannot be
removed by privatizing variables, it is not safe to apply CCO
optimization to the loop.

After applying annotation-based inlining, function calls in the CCO
loop will be eliminated.  Conventional loop dependence analysis can be
applied report the variables that have the loop-carried dependence
discussed in the first section.  Based on the location of the
communication, all the statements in the CCO can be divided into three
groups: (1) computation statements before the communication, (2)
communication statements, and (3) computation statements after the
communication.  Then, the variables that has loop-carried dependence
from (3) to (1) and (2) can be identified using loop dependence
hoisting analysis~\cite{YKA:JSC04}.
